Fixing Vulnerabilities Hinders Maintainability

### 2019-01-30 paper/ICPC19/icpc19.tex
abstract fixed

\begin{abstract}
Security is a requirement of utmost importance to produce software with high
quality. However, there is still a considerable amount of zero-day
vulnerabilities being discovered, and fixed, almost weekly. We suspect that when
developers address these vulnerabilities on their codebases, they actually
affect the maintainability of software, potentially introducing other
vulnerabilities. This paper evaluates the impact of refactorings to improve
security on the maintainability of open-source software. Maintainability is
measured based on the \emph{Better Code Hub}'s model of 10 guidelines using a
dataset including $607$ security commits. Results show that fixing software
vulnerabilities hinders the maintainability of open-source software. This is
particularly the case for refactorings that address \emph{Broken Authentication
and Session Management}, \emph{Memory Leaks} and \emph{Denial-of-Service}
attacks. Therefore, it is necessary to bring awareness for coding practices and
for the use of automated tools that have the potential to help developers with
this issue.
\end{abstract}

@@

Unmaintainable code is hard to test, analyse and re-use.

It is hard to find bugs and vulnerabilities in high complex code, this some developers then to forget maintainability. Therefore, we want to bring awareness for the fact that developers need to write maintainable code to prevent the introduction of vulnerabilities in the code.

The international standard ISO/IEC 25010:2011 breaks down software quality into eight characteristics: maintainability, functional suitability, performance efficiency, compatibility, usability, reliability, security, and portability. This paper focuses solely on maintainability.

In 2011, ISO isued an update for software product quality ISO/IEC 25010~\cite{iso:2011}.

% Our empirical study finds evidence that energy efficiency- oriented code changes have a negative impact on code main- tainability. In particular, careful thinking is required to im- plement the energy patterns Power Save Mode and Wakelock Addition. Furthermore, we show that energy patterns are more likely to require maintenance than regular code changes.

@@

- unsigned long frag_len = msg_hdr->frag_len, max_len; /*!\annotation{lst:func3}!*/

- if((msg_hdr->frag_off+frag_len) > msg_hdr->msg_len)
-    goto err;

@@

\ref{lst:func5} and \ref{lst:func4} add the check to ensure that the message
length in never higher than the maxiumum allowed by the DTSL protocol in
different functions.

@@

@@

\begin{table*}[h]
	\centering
	\caption{Descriptive statistics of the dataset projects}
\begin{tabular}{@{}lllllllllll@{}}
\toprule
      & Forks   & Stars   & Watchers & Contributors & Commits  & Branches & Releases & Size      & Issues & Pull Requests  \\ \midrule
mean  & 1763.52 & 5448.74 & 401.37   & 153.33       & 14834.17 & 45.17    & 129.45   & 122973.24 & 3768.97   & 1941.61 \\
std   & 2434.03 & 6215.09 & 486.60   & 123.68       & 22234.46 & 150.15   & 189.93   & 209732.51 & 5933.16   & 3603.31 \\
min   & 1       & 3       & 1        & 0            & 103      & 1        & 0        & 108       & 0         & 0       \\
25\%  & 391.50  & 1581    & 117.25   & 49           & 1440.50  & 4        & 19       & 8466.75   & 313.75    & 143.25  \\
median  & 838.50  & 2836.50 & 248      & 99           & 5504.50  & 9        & 59       & 37372.50  & 1792.50   & 650     \\
75\%  & 2155    & 6828.50 & 459.50   & 261          & 18579.25 & 20       & 142.75   & 117699.50 & 4087.75   & 1907.25 \\
max   & 16366   & 31841   & 3446     & 413          & 114378   & 1227     & 1114     & 995790    & 33970     & 19329   \\
Total & 165771  & 512182  & 37729    & 14413        & 1394412  & 4246     & 12168    & 11559485  & 354283    & 182511  \\ \bottomrule
\end{tabular}
\label{tab:dataset}
\end{table*}
%
\subsection{Security vs. Baseline Commits}
%
Previous studies attempted to measure the impact of code refactorings on
open-source software maintainability~\cite{HEGEDUS2018313} before. But, to the
best of our knowledge, no previous work has used the BCH model and focused on
evaluating the impact of security refactorings and compare with regular
refactorings as in this study. We analyze the maintainability of regular commits
(i.e., commits not related with security fixes) and use them as a baseline.

@@

\begin{table}[H]
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{2.5cm}L{5.5cm}}

@@

All guidelines are computed in the same evaluation by BCH. During each guideline
evaluation, the tool determines the compliance towards one guideline by
establishing limits for the percentage of code allowed to be in each of the $4$
risk severity levels (\emph{low risk}, \emph{medium risk}, \emph{high risk} and
\emph{very high risk}). If the project does not violate the thresholds, then it
is compliant with the guideline. These thresholds are calibrated by BCH using
their own data/experience --- open-source and closed software systems. If a
project is compliant with a guideline, it means that it is at least $65\%$
better than the software used by BCH to calculate the thresholds\footnote{Check
the answer to \emph{How can I adjust the threshold for passing/not passing a
guideline?} at https://bettercodehub.com/docs/faq (Accessed on \today{})}.
Figure \ref{fig:bchrep} shows an example of the  guidelines report
provided by BCH for a project after finishing its evaluation. These are the
results after performing the OpenSSL CVE-2014-3506 vulnerability refactoring -
motivation example discussed, previously, in section \ref{sec:motivation}. This
version of OpenSSL only complies with 1 out of 10 guidelines, the \emph{Write
Clean Code} guideline.

\begin{figure}[]
 	\centering 	\includegraphics[width=0.45\textwidth]{figures/bch_report.png}
 	\caption{Maintainability report of OpenSSL CVE-2014-3506 vulnerability
refactoring for the guideline \emph{Write Simple Units of Code} provided by
\emph{Better Code Hub}. This version of OpenSSL does not comply with the
guideline in the example since the bars do not reach the threshold points. This
example only complies with $\sfrac{1}{10}$ guidelines (\emph{Write Clean Code}).}
	\label{fig:bchrep}
\end{figure}

@@

\begin{figure}[H]
 	\centering 	\includegraphics[width=0.49\textwidth]{figures/commit.pdf}
 	\caption{Maintainability difference for security commits}
	\label{fig:commit}
\end{figure}

>>>>

\begin{abstract}
Security is a requirement of utmost importance to produce software with high
quality. However, there is still a considerable amount of zero-day
vulnerabilities being discovered, and fixed, almost weekly. We suspect that when
developers address these vulnerabilities on their codebases, they actually
affect the maintainability of software, potentially introducing other
vulnerabilities. This paper evaluates the impact of refactorings to improve
security on the maintainability of open-source software. Maintainability is
measured based on the \emph{Better Code Hub}'s model of 10 guidelines using a
dataset including $607$ security commits. Results show that fixing software
vulnerabilities hinders the maintainability of open-source software. This is
particularly the case for refactorings that address \emph{Broken Authentication
and Session Management} issues, \emph{Memory Leaks} and \emph{Denial-of-Service}
attacks. Within the implications of our project, we highlight the need for tools 
and documentation to help developers adopt security best practices.
\end{abstract}

@@

Unmaintainable code is hard to test, analyse and re-use. It is hard to find bugs and vulnerabilities in high complex code, this some developers then to forget maintainability. Therefore, we want to bring awareness for the fact that developers need to write maintainable code to prevent the introduction of new vulnerabilities in the code or other defects.

Vulnerabilities in open-source software are similar to exploits found in closed products. These are portions of code that were developed with problems or features that attackers can take advantage to carry out harmful actions. In some cases, these issues are the advantage that hackers need to take services down or even get remote access to others systems. However, whereas closed software is written by developers who are following the same organization rules, open-source is less restrictive and more diverse since it depends on the community who writes, fixes, and generally maintains the codebase.

In 2011, the International Organization for Standardization (ISO) issued an update for software product quality ISO/IEC 25010~\cite{iso:2011} considering \emph{Security} as one of the main software product quality characteristics. In this work, we are concerned about the impact that fixing vulnerabilities may have in the maintainability of open-source software. Are developers affecting software maintainability when trying to improve software safety? We suspect that some of those fixes may have a negative impact in the software maintainability and possibly even be the cause of the introduction of new vulnerabilities.

ISO describes software maintainability as “the degree of effectiveness and efficiency with which a software product or system can be modified to improve it, correct it or adapt it to changes in environment, and in requirements” on software quality ISO/IEC 25010. ISO does not provide any specific guidelines or formula to calculate maintainability. Thus, we resort on Software Improvement Group’s web-based source code analysis service Better Code Hub (BCH)~\cite{Visser:2016:OREILLY} to compute the maintainability of a given project. BCH compiles the ISO/IEC 25010 into a set of 10 guidelines that are the base of BCH's evaluation. Two examples of guidelines used by BCH are the \emph{Write Simple Units of Code} based on the McCabe Complexity~\cite{1702388} and \emph{Keep your Codebase Small} based on the idea that smaller codebases are easier to maintain and less sensitive to defects~\cite{Visser:2016:OREILLY}.

In this paper, we present the results of our analysis on the maintainability of 607 security changes collected from open-source software. The main contributions of the present work is an empirical study on the impact of security changes on software maintainability and what patterns need more attention. This empirical study exhibits proof that changes applied in the codebase to fix vulnerabilities affect code maintainability. Especially, for refactorings that address \emph{Broken Authentication and Session Management} issues, \emph{Memory Leaks} and \emph{Denial-of-Service} attacks. With this study, we intend to highlight the need for tools and documentation to help developers 
adopt security best practices.

@@

- unsigned long frag_len = msg_hdr->frag_len, max_len; /*!\annotation{lst:func3}!*/
- if((msg_hdr->frag_off+frag_len) > msg_hdr->msg_len)
-    goto err;

@@

\ref{lst:func5} and \ref{lst:func4} add the check to ensure that the message
length is never higher than the maxiumum allowed by the DTSL protocol in
different functions.

@@

\begin{table*}[h]
	\centering
	\caption{Descriptive statistics of the dataset projects}
\begin{tabular}{@{}lllllllllll@{}}
\toprule
      & Forks   & Stars   & Watchers & Contributors & Commits  & Branches & Releases & Size      & Issues & Pull Requests  \\ \midrule
mean  & 1763.52 & 5448.74 & 401.37   & 153.33       & 14834.17 & 45.17    & 129.45   & 122973.24 & 3768.97   & 1941.61 \\
std   & 2434.03 & 6215.09 & 486.60   & 123.68       & 22234.46 & 150.15   & 189.93   & 209732.51 & 5933.16   & 3603.31 \\
min   & 1       & 3       & 1        & 0            & 103      & 1        & 0        & 108       & 0         & 0       \\
25\%  & 391.50  & 1581    & 117.25   & 49           & 1440.50  & 4        & 19       & 8466.75   & 313.75    & 143.25  \\
median  & 838.50  & 2836.50 & 248      & 99           & 5504.50  & 9        & 59       & 37372.50  & 1792.50   & 650     \\
75\%  & 2155    & 6828.50 & 459.50   & 261          & 18579.25 & 20       & 142.75   & 117699.50 & 4087.75   & 1907.25 \\
max   & 16366   & 31841   & 3446     & 413          & 114378   & 1227     & 1114     & 995790    & 33970     & 19329   \\
Total & 165771  & 512182  & 37729    & 14413        & 1394412  & 4246     & 12168    & 11559485  & 354283    & 182511  \\ \bottomrule
\end{tabular}
\label{tab:dataset}
\end{table*}

@@

%
\subsection{Security vs. Baseline Commits}
%
Previous studies attempted to measure the impact of code refactorings on
open-source software maintainability~\cite{HEGEDUS2018313} before. But, to the
best of our knowledge, no previous work has used the BCH model and focused on
evaluating the impact of security refactorings and compare with regular
refactorings as in this study. We analyze the maintainability of regular commits
(i.e., commits not related with security fixes) and use them as a baseline.

@@

All guidelines are computed in the same evaluation by BCH. During each guideline
evaluation, the tool determines the compliance towards one guideline by
establishing limits for the percentage of code allowed to be in each of the $4$
risk severity levels (\emph{low risk}, \emph{medium risk}, \emph{high risk} and
\emph{very high risk}). If the project does not violate the thresholds, then it
is compliant with the guideline. These thresholds are calibrated by BCH using
their own data/experience --- open-source and closed software systems. If a
project is compliant with a guideline, it means that it is at least $65\%$
better than the software used by BCH to calculate the thresholds\footnote{Check
the answer to \emph{How can I adjust the threshold for passing/not passing a
guideline?} at https://bettercodehub.com/docs/faq (Accessed on \today{})}.

Figure \ref{fig:bchrep} shows an example of the  guidelines report
provided by BCH for a project after finishing its evaluation. These are the
results after performing the OpenSSL CVE-2014-3506 vulnerability refactoring -
motivation example discussed, previously, in section \ref{sec:motivation}. This
version of OpenSSL only complies with 1 out of 10 guidelines, the \emph{Write
Clean Code} guideline.

\begin{table}[h]
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{2.5cm}L{5.5cm}}

@@

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.45\textwidth]{figures/bch_report.png}
 	\caption{Maintainability report of OpenSSL CVE-2014-3506 vulnerability
refactoring for the guideline \emph{Write Simple Units of Code} provided by
\emph{Better Code Hub}. This version of OpenSSL does not comply with the
guideline in the example since the bars do not reach the threshold points. This
example only complies with $\sfrac{1}{10}$ guidelines (\emph{Write Clean Code}).}
	\label{fig:bchrep}
\end{figure}

@@

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.49\textwidth]{figures/commit.pdf}
 	\caption{Maintainability difference for security commits}
	\label{fig:commit}
\end{figure}
